# Tables for Time Series Models

Basic Layout for the individual Time Series Models

- rows = countries
- columns:
  - Coeff with (Std Errors) below and asterisk
  - R2 and adjust R2
  - Newey-West Standard Errors
  - N
  
```{r}
library(tidyverse)
library(sandwich)
library(lmtest)
library(dynlm)
library(modelsummary)
library(gt)
library(ggthemes)
library(patchwork)
```

```{r}
models_B50 <- readRDS("../../data/models/B50_ts_models.rds")
models_M40 <- readRDS("../../data/models/M40_ts_models.rds")
models_T10 <- readRDS("../../data/models/T10_ts_models.rds")
```

Filter EU out

```{r}
models_B50 <- models_B50[names(models_B50) != "EU"]
models_M40 <- models_M40[names(models_M40) != "EU"]
models_T10 <- models_T10[names(models_T10) != "EU"]
```


## Table Creations

```{r}
ts_table <- function(models, table_title) {
  modelsummary(
    models,
    output = "gt",
    vcov = NeweyWest,
    coef_rename = c(
      "d(log(HP_R_N))" = "House Prices",
      "d(log(Stock_Price))" = "Stock Prices"
    ),
    coef_omit = "(Intercept)",
    stars = c("***" = 0.01, "**" = 0.05, "*" = 0.1),
    gof_omit = "IC|RMSE|Log",
    gof_map = list(
      list("raw" = "nobs", "clean" = "N", "fmt" = 0),
      list("raw" = "r.squared", "clean" = "R²", "fmt" = 3),
      list("raw" = "adj.r.squared", "clean" = "Adj. R²", "fmt" = 3)
    ),
    notes = "Standard errors estimated following Newey and West (1987)",
    #shape = model ~ term ,
    title = table_title
  ) %>%
  opt_table_font(size=10)  %>%
  tab_options(latex.tbl.pos = "h")
}
```




```{r}
# Full table 
ts_table_B50 <- ts_table(models_B50, "Bottom 50%: Individual regressions")
gtsave(ts_table_B50, "../../output/ts_tables/ts_B50.html")

# table split up for latex output
ts_table_B50_1 <- ts_table(models_B50[1:11], "Bottom 50%: Individual regressions")
ts_table_B50_2 <- ts_table(models_B50[12:21], NA)

gtsave(ts_table_B50_1, "../../output/ts_tables/ts_B50_1.tex", inline = TRUE)
gtsave(ts_table_B50_2, "../../output/ts_tables/ts_B50_2.tex", inline = TRUE)
```


```{r}
ts_table_M40 <- ts_table(models_M40, "Middle 40%: Individual regressions")
gtsave(ts_table_M40, "../../output/ts_tables/ts_M40.html")

# table split up for latex output
ts_table_M40_1 <- ts_table(models_M40[1:11], "Middle 40%: Individual regressions")
ts_table_M40_2 <- ts_table(models_M40[12:21], NA)

gtsave(ts_table_M40_1, "../../output/ts_tables/ts_M40_1.tex", inline = TRUE)
gtsave(ts_table_M40_2, "../../output/ts_tables/ts_M40_2.tex", inline = TRUE)
```

```{r}
ts_table_T10 <- ts_table(models_T10, "Top 10%: Individual regressions")
gtsave(ts_table_T10, "../../output/ts_tables/ts_T10.html")

# table split up for latex output
ts_table_T10_1 <- ts_table(models_T10[1:11], "Top 10%: Individual regressions")
ts_table_T10_2 <- ts_table(models_T10[12:21], NA)

gtsave(ts_table_T10_1, "../../output/ts_tables/ts_T10_1.tex", inline = TRUE)
gtsave(ts_table_T10_2, "../../output/ts_tables/ts_T10_2.tex", inline = TRUE)
```



## Coefficient Plot

Function to automatically extract the coefficients from the Model

```{r}
make_model_df <- function(models) {
  #generate a df with estimates and std errors of beta 1
  models_df <- data.frame()
  
  for (i in 1:length(models)) {
    model <- models[[i]]
    co <- coeftest(model, vcov = NeweyWest)
    
    estimate_hp = co[2,1] # Coefficient for House Prices
    std_error_hp = co[2,2] # Standard Error for House Prices
    p_value_hp = co[2,4] # p-value for House Prices
    
    estimate_sp = co[3,1] 
    std_error_sp = co[3,2]
    p_value_sp = co[3,4]
    
    models_df <- rbind(models_df, data.frame(
      country = names(models)[i],
      co_hp = estimate_hp,
      se_hp = std_error_hp,
      sign_hp = ifelse(p_value_hp < 0.1,T,F),
      co_sp = estimate_sp,
      se_sp = std_error_sp,
      sign_sp = ifelse(p_value_sp < 0.1,T,F)
    ))
  } 
  models_df
}
```

Function to automatically create the Plots
- creates the error bars
- orders by largest coefficients (not fixed country list!)

```{r}
make_model_plot_hp <- function(models_df) {
  models_df %>%
    ggplot(aes(y = reorder(country, co_hp))) +
    geom_point(aes(x=co_hp, alpha=sign_hp), color="black", size = 2) +
    geom_errorbar(aes(
      xmin = co_hp - 1.645 * se_hp, 
      xmax = co_hp + 1.645 * se_hp, 
      alpha=sign_hp), width = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.4)) +
    labs(
      x = "House Prices Coeff.",
      y = "Country"
    ) +
    theme_tufte() +
    theme(
      legend.position = "none",
      panel.border = element_rect(fill = NA, color = "black"),
    )
    
}

make_model_plot_sp <- function(models_df) {
  models_df %>%
    ggplot(aes(y = reorder(country, co_sp))) +
    geom_point(aes(x=co_sp, alpha=sign_sp), color="black", size = 2) +
    geom_errorbar(aes(
      xmin = co_sp - 1.645 * se_sp, 
      xmax = co_sp + 1.645 * se_sp, 
      alpha=sign_sp), width = 0.2) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.4)) +
    labs(
      x = "Stock Prices Coeff.",
      y = element_blank(),
    ) +
    theme_tufte() +
    theme(
      legend.position = "none",
      panel.border = element_rect(fill = NA, color = "black"),
    )
}
```


```{r}
wrap_coeff_plots <- function(models_df) {
  # automatically create and wrap the plots for the coeffients
  
  p_hp <- make_model_plot_hp(models_df) # housing prices
  p_sp <- make_model_plot_sp(models_df) # stock prices
  
  # wrap the plots together
  wrap_plots(
    p_hp, p_sp, ncol = 2
  ) +
    plot_annotation(
      caption = "Black = significant, Grey = not significant.
      Error bars represent 95% confidence intervals
      Source: ECB DWA, own calculations"
    ) 
}
```


Middle 40% Plot
```{r}
models_df_M40 <- make_model_df(models_M40)
coeff_plot_M40 <- wrap_coeff_plots(models_df_M40)
ggsave("../../output/ts_tables/M40_coeff.png",coeff_plot_M40, width=4, height=5)
# coeff_plot_M40
```


T10 Plot
```{r}
models_df_T10 <- make_model_df(models_T10)
coeff_plot_T10 <- wrap_coeff_plots(models_df_T10)
ggsave("../../output/ts_tables/T10_coeff.png",coeff_plot_T10, width=4, height=5)
#coeff_plot_T10
```


B50 Plot
```{r}
models_df_B50 <- make_model_df(models_B50)
models_df_B50_filter <- models_df_B50 %>% filter(!(country %in% c("NL", "IE")))

coeff_plot_B50 <- wrap_coeff_plots(models_df_B50)
ggsave("../../output/ts_tables/B50_coeff.png",coeff_plot_B50, width=4, height=5)
#coeff_plot_B50
```


Shared House Price Coefficient Plot

```{r}
p_B50_HP <- make_model_plot_hp(models_df_B50) +
# p_B50_HP <- make_model_plot_hp(models_df_B50_filter) +
  labs(title = "Bottom 50%")
  
p_M40_HP <- make_model_plot_hp(models_df_M40) +
  labs(title = "Middle 40%")

p_T10_HP <- make_model_plot_hp(models_df_T10) +
  labs(title = "Top 10%")

p_HP_coeff <- p_T10_HP +
  p_M40_HP +
  p_B50_HP  + 
  plot_layout(ncol = 3, axis_titles = "collect_y") +
  plot_annotation(
    caption = "Black = significant, Grey = not significant.
    Error bars represent 95% confidence intervals
    Source: ECB, BIS, own calculations"
  )
p_HP_coeff

ggsave(p_HP_coeff, file="../../output/ts_tables/HP_coeff.png", width=8, height=4)
# ggsave(p_HP_coeff, file="../../output/ts_tables/HP_coeff_filter.png", width=8, height=4)
```

same for stocks

```{r}
p_B50_SP <- make_model_plot_sp(models_df_B50) +
  labs(title = "Bottom 50%")

p_M40_SP <- make_model_plot_sp(models_df_M40) +
  labs(title = "Middle 40%")

p_T10_SP <- make_model_plot_sp(models_df_T10) +
  labs(title = "Top 10%")

p_SP_coeff <- p_T10_SP +
  p_M40_SP +
  p_B50_SP  + 
  plot_layout(ncol = 3, axis_titles = "collect_y") +
  plot_annotation(
    caption = "Black = significant, Grey = not significant.
    Error bars represent 95% confidence intervals
    Source: ECB, BIS, own calculations"
  )

ggsave(p_SP_coeff, file="../../output/ts_tables/SP_coeff.png", width=8, height=4)
p_SP_coeff

```



