# Counterfactual Simulation

## Mathematical

How would German Top 10 Share have evolved, if the HW was distributed like in EU?

needs
- Net HW of each Decile in 2011
- Overall HW Growth in Germany each quarter
- Share of HW Growth in EU going to each decile


Overall HW Growth EU
$$
\Delta HW_{}^t = HW_{}^{t} - HW_{}^{t-1}
$$

Let Share of HW Growth in EU going to each decile
$$
s_{i}^t = \frac{HW_{i}^{t} - HW_{i}^{t-1}}{\Delta HW_{}^{t}}
$$
with 
$$
\sum_{i=1}^{n} s_i^t = 1
$$



Now Decile growth: 
$$
\Delta HW_{i}^t = s_{i}^t \times \Delta HW_{}^t
$$
leads Decile Level
$$
HW_{i}^{t} = HW_{i}^{t-1} + \Delta HW_{i}^t
$$


The Counterfactual Simulation of Housing Wealth

$$
\hat{HW}_{DE,i}^{t} =  \hat{HW}_{DE,i}^{t-1} + s_{EU,i}^t \times \Delta HW_{DE}^t 
$$
- different from the original growth in Germany (distribution)
- different from the original growth in EU (Delta Growth)

then the counterfactual Total Wealth

$$
\hat{TW}_{DE,i}^{t} = FW_{DE,i}^{t} + BW_{DE,i}^{t} + DEP_{DE,i}^{t} + \hat{HW}_{DE,i}^{t}
$$


## Code

Europe

```{r}
library(tidyverse)
library(ggthemes)
```

```{r}
df <- read_csv("../../data/proc/dwa_simple.csv")
```

### EU shares

```{r}
df_EU <- df %>%
  # filter for the EU
  filter(
    REF_AREA == "EU",
    UNIT_MEASURE == "EUR",
    TIME_PERIOD > "2011-01-01",
  ) %>%
  select(TIME_PERIOD, DWA_GRP, HW_NET) %>% #filter columns
  group_by(TIME_PERIOD) %>%
  mutate(HW_ALL = sum(HW_NET, na.rm = TRUE)) %>% # calculate total HW
  ungroup() %>%
  group_by(DWA_GRP) %>%
  mutate(
    HW_GROWTH_ALL = HW_ALL - lag(HW_ALL, 1), # calculate overall growth
    HW_GROWTH_GRP = HW_NET - lag(HW_NET, 1), # calculate growth per decile
    SHARE_GROWTH = HW_GROWTH_GRP / HW_GROWTH_ALL, # calculate share of growth per decile
  )
```

### German Baseline

```{r}
df_DE <- df %>%
  filter(
    REF_AREA == "DE",
    UNIT_MEASURE == "EUR",
    TIME_PERIOD > "2011-01-01",
  ) %>%
  select(TIME_PERIOD, DWA_GRP, HW_NET) %>%
  group_by(TIME_PERIOD) %>%
  mutate(
    HW_ALL = sum(HW_NET, na.rm = TRUE), # calculate total HW in Germany
  ) %>%
  ungroup() %>%
  group_by(DWA_GRP) %>%
  mutate(
    HW_GROWTH_ALL = HW_ALL - lag(HW_ALL, 1),
    HW_GROWTH_GRP = HW_NET - lag(HW_NET, 1), # calculate growth per decile
  ) 
```

### CTF HW 

combining both DFs

```{r}
df_HW <- df_EU %>%
  select(TIME_PERIOD, DWA_GRP, SHARE_GROWTH) %>%
  right_join(df_DE, by = c("TIME_PERIOD", "DWA_GRP")) %>%
  mutate(
    CTF_HW_GROWTH_GRP = HW_GROWTH_ALL * SHARE_GROWTH, # Counterfactual Growth for each decile
  )
```

Plotting to see if it makes sense

```{r}
df_HW %>%
  filter(DWA_GRP == "B50") %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = HW_GROWTH_GRP, color = "Actual Growth")) +
  geom_line(aes(y = CTF_HW_GROWTH_GRP, color = "Counterfactual Growth")) 
```
now the levels

```{r}
df_HW <- df_HW %>%
  filter(TIME_PERIOD > "2011-04-01") %>% # filter out NA values
  mutate(
    CTF_HW_LEVEL = cumsum(CTF_HW_GROWTH_GRP) + first(HW_NET), # Counterfactual Level for each decile
    HW_LEVEL = cumsum(HW_GROWTH_GRP) + first(HW_NET), # Actual Level for each decile
  )  
```

Robustness check: Totals Add up
```{r}
df_HW %>%
  filter(TIME_PERIOD == "2024-01-01") %>%
  group_by(TIME_PERIOD) %>%
  summarise(
    CTF_HW_LEVEL = sum(CTF_HW_LEVEL, na.rm = TRUE),
    HW_LEVEL = sum(HW_LEVEL, na.rm = TRUE)
  )
```


```{r}
df_HW %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = HW_LEVEL, color = "Actual Level")) +
  geom_line(aes(y = CTF_HW_LEVEL, color = "Counterfactual Level")) +
  facet_wrap(~DWA_GRP, scales = "free_y") 
```

### CTF HW Shares

Create a M40 group
```{r}
df_HW_shares <- df_HW %>%
  mutate(
    DWA_GRP = case_when(
      DWA_GRP %in% c("D06", "D07", "D08", "D09") ~ "M40",
      TRUE ~ DWA_GRP
    )
  ) %>%
  group_by(TIME_PERIOD, DWA_GRP) %>%
  mutate(
    HW_LEVEL = sum(HW_LEVEL, na.rm = TRUE),
    CTF_HW_LEVEL = sum(CTF_HW_LEVEL, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(TIME_PERIOD, DWA_GRP, HW_LEVEL, CTF_HW_LEVEL) %>%
  distinct() # remove duplicates
```
plot levels again with middle class

```{r}
df_HW_shares %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = HW_LEVEL, color = "Actual Level")) +
  geom_line(aes(y = CTF_HW_LEVEL, color = "Counterfactual Level")) +
  facet_wrap(~DWA_GRP, scales = "free_y") 
```



```{r}
df_HW_shares <- df_HW_shares %>%
  group_by(TIME_PERIOD) %>%
  mutate(
    SUM_HW = sum(HW_LEVEL, na.rm = TRUE), # total HW in Germany
    SHARE_HW = HW_LEVEL / SUM_HW, # calculate share of actual HW
    CTF_SHARE_HW = CTF_HW_LEVEL / sum(CTF_HW_LEVEL, na.rm = TRUE) # calculate share of counterfactual HW
  ) 
```

- B50: Share from 0.9 -> 3.2 in 2024
- M40: no change
- D10: Share from 0.55 -> 0.53

= B50 Share would triple, on cost of the top decile

```{r}
df_HW_shares %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = SHARE_HW, color = "Actual Share")) +
  geom_line(aes(y = CTF_SHARE_HW, color = "Counterfactual Share")) +
  facet_wrap(~DWA_GRP, scales = "free_y")
```

### CTF TW

```{r}
df_TW <- df %>%
  filter(
    REF_AREA == "DE",
    UNIT_MEASURE == "EUR",
    TIME_PERIOD > "2011-01-01",
  ) %>%
  # join with calculations of Counterfactuals
  left_join(
    df_HW %>% select(TIME_PERIOD, DWA_GRP, CTF_HW_LEVEL, HW_LEVEL), 
    by = c("TIME_PERIOD", "DWA_GRP")
  ) %>%
  # calculate Total Wealth
  mutate(
    CTF_TW_LEVEL = FW_NET + BW+ DEP + CTF_HW_LEVEL, # Counterfactual Total Wealth
    TW_LEVEL = FW_NET + BW + DEP + HW_LEVEL # Actual Total Wealth
  ) %>%
  select(REF_AREA, TIME_PERIOD, DWA_GRP, CTF_TW_LEVEL, TW_LEVEL, HW_LEVEL, CTF_HW_LEVEL)
```

Plot for DE

```{r}
df_TW %>% 
  filter(
    REF_AREA == "DE"
    ) %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = TW_LEVEL, color = "Actual Total Wealth"), size = 0.5) +
  geom_line(aes(y = CTF_TW_LEVEL, color = "Counterfactual Total Wealth"), size = 0.5) +
  facet_wrap(~DWA_GRP, scales = "free_y") 
  
```

### CTF TW Shares

```{r}
df_TW <- df_TW %>%
  mutate(
    DWA_GRP = case_when(
      DWA_GRP %in% c("D06", "D07", "D08", "D09") ~ "M40",
      TRUE ~ DWA_GRP
    )
  ) %>%
  group_by(TIME_PERIOD, DWA_GRP) %>%
  mutate(
    TW_LEVEL = sum(TW_LEVEL, na.rm = TRUE),
    CTF_TW_LEVEL = sum(CTF_TW_LEVEL, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  select(TIME_PERIOD, DWA_GRP, TW_LEVEL, CTF_TW_LEVEL) %>%
  distinct() # remove duplicates

df_TW_Shares <- df_TW %>%
  group_by(TIME_PERIOD) %>%
  mutate(
    SHARE_TW = TW_LEVEL / sum(TW_LEVEL), # calculate share of actual TW
    CTF_SHARE_TW = CTF_TW_LEVEL / sum(CTF_TW_LEVEL, na.rm = TRUE) # calculate share of counterfactual TW
  )
```

```{r}
df_TW_Shares %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = SHARE_TW, color = "Actual Share"), linewidth = 0.5) +
  geom_line(aes(y = CTF_SHARE_TW, color = "Counterfactual Share"), linewidth = 0.5) +
  facet_wrap(~DWA_GRP, scales = "free_y") 
```


## Fancy Plotting

Bottom 50: CTF Housing Wealth (Level)

```{r}
p1 <- df_HW %>%
  filter(DWA_GRP == "B50") %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = HW_LEVEL, color = "Actual"), linewidth = 0.5) +
  geom_point(aes(y = HW_LEVEL, color = "Actual"), shape = 5) +
  geom_line(aes(y = CTF_HW_LEVEL, color = "Counterfactual"), size = 0.5) +
  geom_point(aes(y = CTF_HW_LEVEL, color = "Counterfactual"), shape = 6) +
  scale_y_continuous(
    labels = scales::label_number(scale = 1e-3),
  ) +
  labs(
    #title = "Simulated Housing Wealth of Bottom 50% in Germany",
    x = "Time Period",
    y = "Housing Wealth (Million EUR)",
    color = ""
  ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black")
  ) 

ggsave("../../output/simulation/B50_HW_LEVEL.png", p1, width = 8, height = 5, dpi = 300)
p1
```

Share of Total Wealth

```{r}
p2 <- df_TW_Shares %>%
  filter(DWA_GRP == "B50") %>%
  ggplot(aes(x = TIME_PERIOD)) +
  geom_line(aes(y = SHARE_TW, color = "Actual"), linewidth = 0.5) +
  geom_point(aes(y = SHARE_TW, color = "Actual"), shape = 5) +
  geom_line(aes(y = CTF_SHARE_TW, color = "Counterfactual"), size = 0.5) +
  geom_point(aes(y = CTF_SHARE_TW, color = "Counterfactual"), shape = 6) +
  labs(
    #title = "Counterfactual Share of Bottom 50% in Germany",
    x = "Time Period",
    y = "Share of Total Wealth",
    color = ""
  ) +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 0.01),
    limits = c(0, 0.04),
    ) +
  theme_tufte() +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, color = "black")
  )
ggsave("../../output/simulation/B50_TW_SHARE.png", p2, width = 8, height = 5, dpi = 300)
p2
```

